---
title: "Lab 02 - Plastic waste"
author: "Truc (Tina) Le"
date: "01/30/22"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
# Load the tidyverse package
library(tidyverse)
```

```{r load-data, message=FALSE}
plastic_waste <-read.csv(here::here("GitHub","ProgDataClass","data","plastic-waste.csv"))
```

## Exercises

### Exercise 1

```{r plastic-waste-continent}
#histogram 

ggplot(data = plastic_waste)+
  aes(x = plastic_waste_per_cap) +
    facet_wrap(~continent)+
  geom_histogram(binwidth = .5)+
  scale_x_continuous(name = "Plastic Waste per Capita")

#density
ggplot(data = plastic_waste)+
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent)+
  geom_density()+
   scale_x_continuous(name = "Plastic Waste per Capita")

#history and density
ggplot(data = plastic_waste)+
  aes(x = plastic_waste_per_cap, 
      fill = continent,
      color = continent) +
  geom_histogram(binwidth = .5)+
  facet_wrap(~continent)+
  geom_density(aes(y = after_stat(count)),
               alpha = .1)+
   scale_x_continuous(name = "Plastic Waste per Capita")

#Fill and color were defined by the mapping aesthetic or aes() because we wanted to apply the color to both the graph and the chart. The alpha level was directly inputted into the geom since we only wanted to change the opacity for the density graph.
 
  
```

### Exercise 2

```{r plastic-waste-density}
# boxplot
ggplot(data = plastic_waste)+
  aes(x = continent,
      y = plastic_waste_per_cap,
      fill = continent)+
  geom_boxplot()

# box plot w/ individual points

ggplot (data = plastic_waste)+
  aes(x = continent,
      y = plastic_waste_per_cap,
      fill = continent,
      color = continent)+
  geom_boxplot(color = "black")+
  geom_jitter(height = .5,
              alpha = .3)

#boxplot w/density curve "raincloud plot"

ggplot(data = plastic_waste)+
  aes(x = continent,
      y = plastic_waste_per_cap,
      fill = continent,
      color = continent)+
  geom_jitter(height = .2)+
   geom_boxplot(color = "black",
                alpha = .5,
                width = 1,
                size = .5)+
  ggdist::stat_slab (height = .5,
                     color = "black",
                     size = .2,
                     alpha = .5,
                     position = position_nudge (y = .2))
```

### Exercise 3

```{plastic-waste-per-capita}

```{r plastic-waste-per-capita}
#scatterplot

ggplot(data = plastic_waste)+
  aes(x = plastic_waste_per_cap,
  y = mismanaged_plastic_waste_per_cap,
  )+
   geom_point()+
   scale_x_continuous(name = "Plastic Waste per Capita")+
   scale_y_continuous(name = "Mismanaged Plastic Waste per Capita")

#There is an upward trend and positive relationship between plastic waste per capita and mismanaged plastic waste per capita.

#scatterplot with color

ggplot(data = plastic_waste)+
  aes(x = plastic_waste_per_cap,
  y = mismanaged_plastic_waste_per_cap,
  color = continent)+
   geom_point()+
   scale_x_continuous(name = "Plastic Waste per Capita")+
   scale_y_continuous(name = "Mismanaged Plastic Waste per Capita")

#Europe has lower plastic waste per capita and mismanaged plastic waste per capita than the other continents. Asia has the highest mismanaged plastic waste per capita while North America has the highest plastic waste per capita.

#Plastic Waste per Capita and Total Population

ggplot(data = plastic_waste) +
  aes(x = total_pop,
      y = plastic_waste_per_cap) +
  geom_point() +
  geom_smooth()

#Plastic Waste per Capita and Coastal Population

ggplot(data = plastic_waste) +
  aes(x = coastal_pop,
      y = plastic_waste_per_cap) +
  geom_point() +
  geom_smooth()

#Plastic waste per capita and coastal population has the stronger relationship with less variability
```

### Bonus

Remove this text, and add your answer for Exercise 4 here.


.
